---
title: "Introduction to CorporaCoCo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to CorporaCoCo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#",
    error = FALSE,
     tidy = FALSE,
    cache = FALSE,
 collapse = TRUE)
options(width = 100)
```
# Getting started

Load the library.
This is available from [CRAN](https://cran.r-project.org/package=CorporaCoCo) and the latest development versions can be found on [GitHub](https://github.com/mahlberg-lab/CorporaCoCo).
```{r}
library(CorporaCoCo)
```

## corp\_text

A passage from *Alice's Adventures in Wonderland*.
```{r}
a <- "`But do cats eat bats, I wonder?' And here Alice began to get rather sleepy, and went on saying to herself, in a dreamy sort of way, `Do cats eat bats? Do cats eat bats?' and sometimes, `Do bats eat cats?' for, you see, as she couldn't answer either question, it didn't much matter which way she put it."
```
Create a `corp_text` object
```{r}
a_text <- corp_text(a)
```
We used the default tokenization, so lets look at the type mappings:
```{r}
corp_type_lookup(a_text)
```
You can combine `corp_text` objects
```{r}
aa <- "The cat sat on the mat."
bb <- "This dog ate this cat."
aaa <- corp_text(aa)
bbb <- corp_text(bb)
ccc <- list(aaa, bbb)
corp_text_rbindlist(ccc)

```

## corp\_concordance
You can look at concordance lines using `corp_text` objects.
```{r}
corp_concordance(a_text, span = "4LR", nodes = c("bats", "cats"))
```
If you supply `collocates` they will be highlighted (they will not act as a filter)
```{r}
y <- corp_concordance(a_text, span = "4LR", nodes = c("bats", "cats"), collocates = c("sometimes", "eat"))
y
```

Because a `corp_concordance` object is just a `data.table` you can sort it, filter it, save it to a CSV file just like a `data.table`
```{r}
y[order(N, R1_type)]
```
and you can do this using the *types* or the *tokens*
```{r}
names(y)
```
and if you want to see the raw `data.table`.
```{r}
print(y, as_data_table = TRUE)
```

## corp\_surface
This works just like the old `surface` function but now you must pass it `corp_text` objects.
```{r}
a_cooccurs <- corp_surface(a_text, span = '1L1R', nodes = c("alice", "bats", "cats"))
```
BTW if you do `corp_concordance` on a `corp_surface` object it will use the `span`, `nodes` and `collocates` values used to create the `corp_surface` object as the default values for `corp_concordance`
```{r}
corp_concordance(a_cooccurs)
```

## corp\_coco
The same passage after it has been translated into Finnish and back to english using Google translate.
```{r}
b <- "`But cats eat bats, I wonder?' And here, Alice began to get pretty sleepy and went on to say to herself, in a dreamlike way: `Are cats eating bats? Are cats eating bats?' And sometimes, `Do cats have to eat cats?' since you see because he could not answer the questions, he did not really matter how he put it."
b_text <- corp_text(b)

b_cooccurs <- corp_surface(b_text, span = '1L1R', nodes = c("alice", "bats", "cats"))
b_cooccurs
```

```{r}
cats_and_bats <- corp_coco(a_cooccurs, b_cooccurs, nodes = c("alice", "bats", "cats"), fdr = 1.0)
plot(cats_and_bats)
```

## Complex types

Looks like their is a difference for *cats* and *bats* but the text is small so we have no power.
Lets have a look at the tokenization.

```{r}
corp_type_lookup(a_text)
```

Since *cats* and *bats* are both types of mammals maybe we can look at the co-occurences with *mammals*.
```{r}
a_tokens <- corp_get_tokens(a_text)
b_tokens <- corp_get_tokens(b_text)

a_tokens$type <- sub("^(cats|bats)$", "MAMMALS", a_tokens$type)
b_tokens$type <- sub("^(cats|bats)$", "MAMMALS", b_tokens$type)

a_text_mammals <- corp_text(a, tokens = a_tokens)
b_text_mammals <- corp_text(b, tokens = b_tokens)
```
So now we have a new type to token mapping
```{r}
corp_type_lookup(a_text_mammals)

```
and so
```{r}
a_cooccurs_mammals <- corp_surface(a_text_mammals, span = '1L1R', nodes = c("alice", "MAMMALS"))
b_cooccurs_mammals <- corp_surface(b_text_mammals, span = '1L1R', nodes = c("alice", "MAMMALS"))

mammals <- corp_coco(a_cooccurs_mammals, b_cooccurs_mammals, nodes = c("alice", "MAMMALS"), fdr = 1.0)
plot(mammals)

```



