---
title: "Introduction to CorporaCoCo"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    css: styles.css
vignette: >
  %\VignetteIndexEntry{Introduction to CorporaCoCo}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#",
    error = FALSE,
     tidy = FALSE,
    cache = FALSE,
 collapse = TRUE)
options(width = 100)
```
# Getting started

Load the library.
This is available from [CRAN](https://cran.r-project.org/package=CorporaCoCo) and the latest development versions can be found on [GitHub](https://github.com/birmingham-ccr/CorporaCoCo).
```{r}
library(CorporaCoCo)
```

## Concordance

A passage from *Alice's Adventures in Wonderland*.
```{r}
a <- "`But do cats eat bats, I wonder?' And here Alice began to get rather sleepy, and went on saying to herself, in a dreamy sort of way, `Do cats eat bats? Do cats eat bats?' and sometimes, `Do bats eat cats?' for, you see, as she couldn't answer either question, it didn't much matter which way she put it."
```
Create a co-occurrence object
```{r}
x <- corp_surface(a, span = "4LR", nodes = "eat", collocates = c("bats", "cats"))
```
You can then use this object to examine the concordance for a set of node types
```{r}
corp_concordance(x)
```
Note that it has used the `span`, `nodes` and `collocates` that you supplied to `corp_surface()`.

Actually you can supply these directly to `corp_concordance()`
```{r}
y <- corp_concordance(x, span = "2L4R", nodes = c("cats", "bats"), collocates = "eat") 
y
```
Because a `corp_concordance` object is just a `data.table` you can sort it, filter it, save it to a CSV file just like a `data.table`
```{r}
y[order(N, R1)]
```
and you can do this using the *types* or the *tokens*
```{r}
names(y)
```

## Complex types

A passage from *Alice's Adventures in Wonderland*.
```{r}
a <- "`But do cats eat bats, I wonder?' And here Alice began to get rather sleepy, and went on saying to herself, in a dreamy sort of way, `Do cats eat bats? Do cats eat bats?' and sometimes, `Do bats eat cats?' for, you see, as she couldn't answer either question, it didn't much matter which way she put it."

a_cooccurs <- corp_surface(a, span = '1L1R', nodes = c("alice", "bats", "cats"))
a_cooccurs
```

The same passage after it has been translated into Finnish and back to english using Google translate.
```{r}
b <- "`But cats eat bats, I wonder?' And here, Alice began to get pretty sleepy and went on to say to herself, in a dreamlike way: `Are cats eating bats? Are cats eating bats?' And sometimes, `Do cats have to eat cats?' since you see because he could not answer the questions, he did not really matter how he put it."

b_cooccurs <- corp_surface(b, span = '1L1R', nodes = c("alice", "bats", "cats"))
b_cooccurs
```

```{r}
cats_and_bats <- corp_coco(a_cooccurs, b_cooccurs, nodes = c("alice", "bats", "cats"), fdr = 1.0)
plot(cats_and_bats)
```

Looks like their is a difference for *cats* and *bats* but the text is small so we have no power.
Lets have a look at the tokenization.

```{r}
corp_type_lookup(a_cooccurs)
```

Since *cats* and *bats* are both types of mammals maybe we can look at the co-occurences with *mammals*.
```{r}
a_tokens <- corp_get_tokens(a_cooccurs)
b_tokens <- corp_get_tokens(b_cooccurs)

a_tokens$type <- sub("^(cats|bats)$", "MAMMALS", a_tokens$type)
b_tokens$type <- sub("^(cats|bats)$", "MAMMALS", b_tokens$type)

a_cooccurs_mammals <- corp_surface(a, tokens = a_tokens, span = '1L1R', nodes = c("alice", "MAMMALS"))
b_cooccurs_mammals <- corp_surface(b, tokens = b_tokens, span = '1L1R', nodes = c("alice", "MAMMALS"))
```

So now we have a new type to token mapping
```{r}
corp_type_lookup(a_cooccurs_mammals)

```
and so
```{r}
mammals <- corp_coco(a_cooccurs_mammals, b_cooccurs_mammals, nodes = c("alice", "MAMMALS"), fdr = 1.0)
plot(mammals)

```



